Declarative runtime reconfiguration of pulverised distributed systems \thanks{Identify applicable funding agency here. If none, delete this.}

### 2023-05-12 paper-2023-acsos-pulverization.tex
Meta with limit and deadline
More related work

@@

This work provides 
 \acp{dsl}
 for specifying 
 distributed systems and
 corresponding deployment plans,
 and a framework for executing
 deployment plans and reconfigurations.
%
Therefore, we provide an original solution
 to different issues.
%
In~\cite{DBLP:journals/jss/ArcangeliBL15},
the problem of (automatic) deployment of distributed software system is studied by considering
 (i) the nature of software to be deployed,
 (ii) the nature of the deployment domain (i.e. the available infrastructure),
 (iii) how the deployment is designed,
 and 
 (iv) how the deployment is performed.
%
Accordingly, 
 we cover related works about
 languages for 
 specifying distributed systems (\Cref{sec:rw:whatsw}),
 languages for specifying infrastructures and deployments (\Cref{sec:rw:depdesc}),
 and
 techniques for automatic deployment and reconfiguration (\Cref{sec:rw:autodep}).

@@

Specifically,
 this work generalises over the \emph{pulverisation model}~\cite{FI2020-pulverization,IEEE-IoTJ-pulverization-simulation},
 whereby a large-scale cyber-physical system
 is partitioned into a graph of devices
 where each device is split into five deployable components:
 (i) sensors interface, 
 (ii) actuators interface, 
 (iii) behaviour, 
 (iv) state, and 
 (v) communication component.
%
In~\cite{IEEE-IoTJ-pulverization-simulation},
 a methodology on top of the pulverisation model
 is proposed
 for generating and testing deployments through simulation.

Also,
 the approach of specifying in a single codebase 
 parts of the structure, behaviour, and/or interaction 
 for a whole distributed system
 is also related to \emph{macroprogramming}~\cite{Casadei2023macro}
 and, in particular, to \emph{multi-tier programming} paradigm~\cite{DBLP:journals/csur/WeisenburgerWS20},
 where the deployment units for different tiers (e.g., client and server tiers; or view, business logic, and data tiers) are obrained by compilation or interpretation of a single codebase.

@@

\subsection{Automatic deployment and reconfiguration}
\label{sec:rw:autodep}
%
A comprehensive conceptual framework
 on automatic deployment of distributed systems
 is provided in a 2015 survey by Arcangeli et al. \cite{DBLP:journals/jss/ArcangeliBL15}.
%
A more recent survey by Coullon et al. focusses
 on formal techniques for verifying the correctness of reconfigurations
 in component-based distributed software systems~\cite{coullon2023swreconfig}.
%
\meta{TODO: say something about our positioning wrt verification of reconfigurations}

@@

\meta{
MuScADeL (Multi-Scale Autonomic Deployment Language)~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}

ADSL (AWaRE DSL)~\cite{DBLP:conf/aswec/ChhetriLUVKNR18/adsl} % and see corresp RW: the most related DSLs are those in the area of automated deployment [17]. Such DSLs include Deladas [31], [32], ConfSolve [15], [33], j-ASD’s DSL [34] and MuScADeL [35], [36].
% [17]: J.-P. Arcangeli, R. Boujbel, and S. Leriche, "Automatic deployment of distributed software systems: Definitions and state of the art," Journal of Systems and Software, vol. 103, pp. 198–218, 2015

%
%\subsection{Approaches to Deployment Independence}
%\label{s:rw:deployment-independence}
%
%\meta{TODO: add works on deployment independence, e.g., osmotic~\citep{DBLP:journals/computer/VillariFDRJR19}, BIP~\citep{lekidis2015bip-wsn,bastarrica2001optimization-techniques-deployment-components-bip}}
%
%We highlight a number of representatieve research efforts on deployment and automatic reconfiguration of systems.
%%
%%For instance, 
%\emph{Osmotic computing}~\cite{DBLP:journals/computer/VillariFDRJR19} is an approach to opportunistic deployment of microservices on the edge-fog-cloud platform.
%An osmotic platform aims to reach and maintain an ``osmotic equilibrium'' 
%between infrastructural and application requirements
%by automatically migrating microservices to deployment locations.
%
However, the approach mainly targets centrally orchestrated systems.
%
Other approaches leverage component-based, architectural descriptions to decouple application logic and deployment.
%
For instance, \emph{DR-BIP (Dynamically Reconfigurable - Behaviour Interaction Priority model)}~\cite{,Ballouli18dr-bip}
and \emph{DReAM (Dynamically Reconfigurable Architectural Modelling)}~\cite{denicola2020dream-dynamic-reconfig-arch-modelling}
use \emph{components} (capturing behaviour), \emph{connectors} (capturing interaction between components' \emph{ports}), 
\emph{maps} (logical topologies),
and \emph{deployments} (associating components to map locations),
overall organised in \emph{motifs} (dynamic architectural configurations),
to model and analyse dynamic architectures.
%

These approaches have some similarity with the approach presented in this paper, but they are arguably more complex and our  
%than the one introduced in this paper,
 %, though also fostering an exogenous, declarative (logics-based) modelling approach.
approach explicitly addresses self-organising CPS.
%
%
%By constrast, our approach explicitly addresses self-organising CPS.
%
%A more in-depth, formal comparison is left as future work.
}

>>>>

\meta{LIMIT 10 pages incl. refs; DEADLINE 19 may}

@@

This work provides 
 \acp{dsl}
 for specifying 
 distributed systems (partitioned according to the 
 pulverisation model)
 and
 corresponding deployment plans,
 and a framework for executing
 deployment plans and reconfigurations.
%
A way to structure all this 
 is by using the analytical framework of~\cite{DBLP:journals/jss/ArcangeliBL15},
  where the problem of (automatic) deployment of distributed software system is studied by considering
 (i) the nature of software to be deployed,
 (ii) the nature of the deployment domain (i.e. the available infrastructure),
 (iii) how the deployment is designed,
 and 
 (iv) how the deployment is performed.
%
Accordingly, 
 we cover related works about
 languages for 
 specifying distributed systems (\Cref{sec:rw:whatsw}),
 languages for specifying infrastructures and deployments (\Cref{sec:rw:depdesc}),
 and
 techniques for automatic deployment and reconfiguration (\Cref{sec:rw:autodep}).

@@

Specifically,
 this work builds on the \emph{pulverisation} component model~\cite{FI2020-pulverization,IEEE-IoTJ-pulverization-simulation},
 whereby a large-scale cyber-physical system
 is partitioned into a graph of devices
 where each device is split into five deployable components:
 (i) sensors interface, 
 (ii) actuators interface, 
 (iii) behaviour, 
 (iv) state, and 
 (v) communication component.
%
In~\cite{IEEE-IoTJ-pulverization-simulation},
 a methodology on top of the pulverisation model
 is proposed
 for generating and testing deployments through simulation.

Additionally,
 the approach of specifying in a single codebase 
 parts of the structure, behaviour, and/or interaction 
 for a whole distributed system
 is also related to \emph{macroprogramming}~\cite{Casadei2023macro}
 and, in particular, to \emph{multi-tier programming} paradigm~\cite{DBLP:journals/csur/WeisenburgerWS20},
 where the deployment units for different tiers (e.g., client and server tiers; or view, business logic, and data tiers) are obrained by compilation or interpretation of a single codebase.

@@

A \emph{deployment plan} is a configured mapping of a software system onto a deployment domain.
%
Given a deployable software system developed using the techniques of the previous subsection,
 what is needed now is a language to describe a deployment domain 
 and the deployment mapping.
%
Languages have been proposed 
 targetting specific kinds of deployment domains
 such as the smart grid (cf. the deployment specification language \emph{dspec} in the \emph{Resilient Information Architecture Platform for Smart Grids [RIAPS]} platform~\cite{DBLP:conf/coins/GhoshTKKL22})
 or the cloud (cf. the \emph{Cloud Application Modelling and Execution Language [CAMEL]} multi-\ac{dsl}~\cite{DBLP:journals/jcloudc/AchilleosKRKDOS19}).

In this work,
 we are especially interested deployments over large-scale cyber-physical systems and the edge-fog-cloud continuum.
%
The \emph{MultiScale Autonomic Deployment Language (MuScADeL)}~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}
 is a \ac{dsl} that allows to express deployment properties of 
 componentised applications on multi-scale deployment domains, i.e., large and heterogeneous infrastructures considered under several viewpoints (device, geography, network, administrative domains, etc.).
%
In~\cite{DBLP:journals/sosym/SongDFSF22},
  a model-based deployment approach is proposed
  targetting so-called \emph{fleets}, i.e.,
  distributed and heterogeneous devices at the edge chaaracterised by different cyber-physical contexts (cf. resources, connectivity, etc.).
%
It is based on the \emph{Generation and Deployment
of Smart IoT Systems (GeneSIS)} modelling language, where a \texttt{DeploymentModel} consists of sets of \texttt{Resources} (e.g., \texttt{Component}s, specialised by \texttt{InfrastructureComponent}s and \texttt{SoftwareComponent}s) with associated \texttt{Property}s and \texttt{Link}s.

\subsection{Automatic deployment and reconfiguration}
\label{sec:rw:autodep}
%
A comprehensive conceptual framework
 on automatic deployment of distributed systems
 is provided in a 2015 survey by Arcangeli et al. \cite{DBLP:journals/jss/ArcangeliBL15}.
%
A more recent survey by Coullon et al. focusses
 on formal techniques for verifying the correctness of reconfigurations
 in component-based distributed software systems~\cite{coullon2023swreconfig}.
%
Two formal models for specifying reconfigurable architectures are \emph{DR-BIP (Dynamically Reconfigurable - Behaviour Interaction Priority model)}~\cite{,Ballouli18dr-bip}
and \emph{DReAM (Dynamically Reconfigurable Architectural Modelling)}~\cite{denicola2020dream-dynamic-reconfig-arch-modelling}.
%
Both are based on the same conceptual model: they 
 feature \emph{components} (capturing behaviour), \emph{connectors} (capturing interaction between components' \emph{ports}), 
\emph{maps} (logical topologies),
and \emph{deployments} (associating components to map locations),
overall organised in \emph{motifs} (dynamic architectural configurations),
to model and analyse dynamic architectures.
%
\meta{TODO: say something about our positioning wrt verification of reconfigurations}

@@

%\meta{
%MuScADeL (Multi-Scale Autonomic Deployment Language)~\cite{DBLP:conf/compsac/BoujbelRLTAL14/muscadel}
%
%ADSL (AWaRE DSL)~\cite{DBLP:conf/aswec/ChhetriLUVKNR18/adsl} % and see corresp RW: the most related DSLs are those in the area of automated deployment [17]. Such DSLs include Deladas [31], [32], ConfSolve [15], [33], j-ASD’s DSL [34] and MuScADeL [35], [36].
%% [17]: J.-P. Arcangeli, R. Boujbel, and S. Leriche, "Automatic deployment of distributed software systems: Definitions and state of the art," Journal of Systems and Software, vol. 103, pp. 198–218, 2015
%%
%%%
%%%\subsection{Approaches to Deployment Independence}
%%%\label{s:rw:deployment-independence}
%%%
%%%\meta{TODO: add works on deployment independence, e.g., osmotic~\citep{DBLP:journals/computer/VillariFDRJR19}, BIP~\citep{lekidis2015bip-wsn,bastarrica2001optimization-techniques-deployment-components-bip}}
%%%
%%%We highlight a number of representatieve research efforts on deployment and automatic reconfiguration of systems.
%%%%
%%%%For instance, 
%%%\emph{Osmotic computing}~\cite{DBLP:journals/computer/VillariFDRJR19} is an approach to opportunistic deployment of microservices on the edge-fog-cloud platform.
%%%An osmotic platform aims to reach and maintain an ``osmotic equilibrium'' 
%%%between infrastructural and application requirements
%%%by automatically migrating microservices to deployment locations.
%%%
%%However, the approach mainly targets centrally orchestrated systems.
%%%
%%Other approaches leverage component-based, architectural descriptions to decouple application logic and deployment.
%%%
%%For instance, \emph{DR-BIP (Dynamically Reconfigurable - Behaviour Interaction Priority model)}~\cite{,Ballouli18dr-bip}
%%and \emph{DReAM (Dynamically Reconfigurable Architectural Modelling)}~\cite{denicola2020dream-dynamic-reconfig-arch-modelling}
%%use \emph{components} (capturing behaviour), \emph{connectors} (capturing interaction between components' \emph{ports}), 
%%\emph{maps} (logical topologies),
%%and \emph{deployments} (associating components to map locations),
%%overall organised in \emph{motifs} (dynamic architectural configurations),
%%to model and analyse dynamic architectures.
%%%
%%
%%These approaches have some similarity with the approach presented in this paper, but they are arguably more complex and our  
%%%than the one introduced in this paper,
%% %, though also fostering an exogenous, declarative (logics-based) modelling approach.
%%approach explicitly addresses self-organising CPS.
%%%
%%%
%%%By constrast, our approach explicitly addresses self-organising CPS.
%%%
%%%A more in-depth, formal comparison is left as future work.
%}
